Constructing the transition graphs of interacting hysterons

### 2022-06-03 Paper_draft.tex
Minor change: 'in previous work from our group' replaced by 'in recent work
Reworked section on avalanches by expanding on how to construct l=2 avalanches & added corresponding figure

In past work from our group, the possible transition graphs for two- and three-hysteron systems were found using numerical sampling. The obtained graphs contain several noteworthy phenomena that are not possible in the Preisach model, such as breaking of loop return point memory, scrambling, multiperiodic cycles and long transients (Van Hecke, 2021).\\

@@

    \includegraphics[width=0.5\textwidth]{constructing_avalanches_sketch.png}
    \caption{Construction of an example $N=2$ graph by iteratively building up the avalanche $11\downarrow 10\downarrow 00$.}

@@

We here show how the possible avalanches are evaluated for a given base graph.\\
We first note that we can build up avalanches iteratively. Namely, to have the $l=3$ avalanche $S^{(0)}\rightarrow S^{(1)} \rightarrow S^{(2)}$, one must at least have the $l=2$ avalanche $S^{(0)}\rightarrow S^{(1)}$. Thus, we start by evaluating all $l=2$ avalanches.\\
To have a possible avalanche $S^{(0)}\rightarrow S^{(1)} \rightarrow S^{(2)}$, there must be an incoming transition $S^{(0)}\rightarrow S^{(1)}$ at the state $S^{(1)}$, as well as an outcoming transition $S^{(1)} \rightarrow S^{(2)}$. Therefore, we can systematically find all possible $l=2$ avalanches in a base graph by constructing all combinations of incoming and outgoing transitions at each state. In addition, it must be checked for each avalanche that $S^{(0)} \neq S^{(2)}$, as an avalanche cannot revisit the same state.\\
Next, we consider how the avalanche length $l$ can be iterated over to obtain graphs with longer avalanches. In the previous step, we have explored all possible combinations of $l=2$ avalanches for the given base graph. To have an $l=3$ avalanche, an $l=2$ avalanche must be extended with an additional transition. Thus, similarly to our method for finding the $l=2$ avalanches, we find all $l=3$ avalanches for a given graph with only $l=2$ by evaluating all combinations of incoming $l=2$ and outgoing $l=1$ transitions at each state.\\
In summary, to construct all possible graphs for $N$ hysterons, we first construct all base graphs. We then generate the candidate graphs for each base graph by constructing the possible avalanches, and evaluating each combination of avalanches.\\
\section{Application to specific models}
We now illustrate how the base graph method is applied to the specific model where the coupling between hysterons is linear. Linear coupling entails that the switching fields $U_i^{+, -}(S)$ depend on the state $S$ as:
    U_i^{+, -}(S) = u_i^{+, -} -\sum_{i \neq j}c_{ij}s_j
Depending on the system, the allowed coupling coefficients $c_{ij}$ may be restricted. We consider two instances of the linear model for $N=3$. First, we look at the case where only one of the coupling coefficients $c_{ij}$ is unequal to zero. Second, we look at a model for hysterons in series, where the coupling is $c_{ij}  = c_j$ with all $c_j < 0$.\\
\subsection{Linear coupling}
%\begin{figure}[!ht]
%    \centering
%    \includegraphics[width=0.5\textwidth]{.png}
%    \caption{Overview of possible scrambling and avalanches in the linear model for a single nonzero $c_ij$.}
%    \label{fig:linear_coupling}
%\end{figure}\\
\subsection{Hysterons in series}
\section{Discussion}
\section{Outlook}
%\bibliography{apssamp}% Produces the bibliography via BibTeX.

>>>>

In recent work, the transition graphs for two- and three-hysteron systems were found using numerical sampling. The obtained graphs contain several noteworthy phenomena that are not possible in the Preisach model, such as breaking of loop return point memory, scrambling, multiperiodic cycles and long transients (Van Hecke, 2021).\\

@@

    \includegraphics[width=0.5\textwidth]{constructing_avalanches_sketch_2.png}
    \caption{Construction of an $l=2$ avalanche by combining an incoming transition with one of at most two outgoing transitions. a) Combination of an incoming up transition with an outgoing up transition (ferromagnetic) or down transition (antiferromagnetic). b) Combination of an incoming down transition with an outgoing up transition (antiferromagnetic) or down transition (ferromagnetic) }

@@

We here show how the possible avalanches are evaluated for a given base graph.We first describe how one systematically constructs the possible avalanches of length $l=2$ for a base graph. We then generalize our method to avalanches of any length $l$\\
Let a base graph contain the up transition $S^{(0)}\uparrow S^{(1)}$. The state $S^{(1)}$ has two possible transitions: an up transition $S^{(1)}\uparrow S^{(2+)}$, and a down transition  $S^{(1)}\downarrow S^{2-}$.  The transition $S^{(0)}\uparrow S^{(1)}$ can thus be extended to a $l=2$ avalanche in two ways: either by the ferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\uparrow S^{(2+)}$, or by the antiferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}a). Similarly, a down transition $S^{(0)}\downarrow S^{(1)}$ can be extended to the antiferromagnetic avalanche $S^{(0)}\downarrow S^{(1)}\uparrow S^{(2+)}$, or to the ferromagnetic avalanche  $S^{(0)}\downarrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}b). Therefore, for each transition $S^{(0)}\rightarrow S^{(1)}$, there is one possible ferromagnetic and one antiferromagnetic avalanche when $S^{(1)}$ is not a saturated state. In the case where $S^{(1)}$ is saturated, only the antiferromagnetic option exists.\\
 We see that every transition $S^{(0)}\rightarrow S^{(1)}$ in the base graph can be extended to a $l=2$ transition  $S^{(0)}\rightarrow S^{(1)}\rightarrow S^{(2)}$ in at most two ways. However, not every avalanche is necessarily valid. This is because an avalanche may not contain any cycles: if an avalanche did contain a cycle, it would lead to an infinite loop. We must therefore impose that $S^{(2)} \neq S^{(0)}$. As a result, the number of avalanche options per transition is either two, one, or zero.\\

\section{Application to specific models}
We now illustrate how the base graph method is applied to the specific model where the coupling between hysterons is linear. Linear coupling entails that the switching fields $U_i^{+, -}(S)$ depend on the state $S$ as:
\begin{equation}
    U_i^{+, -}(S) = u_i^{+, -} -\sum_{i \neq j}c_{ij}s_j
\end{equation}
Depending on the system, the allowed coupling coefficients $c_{ij}$ may be restricted. We consider two instances of the linear model for $N=3$. First, we look at the case where only one of the coupling coefficients $c_{ij}$ is unequal to zero. Second, we look at a model for hysterons in series, where the coupling is $c_{ij}  = c_j$ with all $c_j < 0$.\\
\subsection{Linear coupling}
%\begin{figure}[!ht]
%    \centering
%    \includegraphics[width=0.5\textwidth]{.png}
%    \caption{Overview of possible scrambling and avalanches in the linear model for a single nonzero $c_ij$.}
%    \label{fig:linear_coupling}
%\end{figure}\\
\subsection{Hysterons in series}
\section{Discussion}
\section{Outlook}
%\bibliography{apssamp}% Produces the bibliography via BibTeX.
